---
title: "Project 1"
author: "Isaac"
format: html
editor: visual
---

## Querying Census API

The goal of this project is to query the "Public Use Microdata Sample" (PUMS) API from the Census Bureau. We do this by first building the URL to contact a particular endpoint of the PUMS API. In particular, we will go with the example call provided in the PUMS API under their most recent year, 2022, which we store as a string.

```{r}
URL_pums <- "https://api.census.gov/data/2022/acs/acs1/pums/variables.json"
```

Next we use the `httr::GET()` function to contact the website and explore its structure.

```{r}
pums_info <- httr::GET(URL_pums)
str(pums_info, max.level = 1)
```

This is a lot of information and not very clear or readable, so we will use the `jsonlite` package to parse this data.

```{r}
library(jsonlite)

#parse data using rawToChar() function 
parsed <- fromJSON(rawToChar(pums_info$content))

#turn parsed data into nice tibble
acs_estimates_info <- tibble::as_tibble(parsed$data)

#display tibble
acs_estimates_info
```
